Developing a preprint culture in biology

### 2012-09-28 ms.tex
comment

While submitting to public preprints servers is still uncommon in biology,
preprints are becoming more common in biological sciences. The quantitative
biology section in arXiv is experiencing faster growth in submissions than any
other field \cite{cal12}. Also, most scientific journals are preprint-friendly:
Nature, PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the
journals from Elsevier and Springer. Very recently, the Ecological Society of
America and the Genetics Society of America changed their policies to allow
public preprints. Few scientific publications will not consider a manuscript
submitted to a public preprint server.  Still, a few journals adopt a ``by
default'' hostile attitude towards preprints, mostly due to the lack of clear
policy of the publishers, or perhaps because a preprint culture has not
developed in biology. As an example, Wiley-Blackwell, which publishes some of
the leading journal in the field, has no official policy on the subject
\ref{table:policies}.

@@

There are other barriers to adoption of preprint servers by biologists.  The
most notable preprint server, arXiv, is designed to work best with \LaTeX, the
dominant document preparation system among physicists and mathematicians.
Jackson \cite{jackson2002preprints} further argues that \LaTeX introduced an
open source mindset to its users, who now freely share their research
findings as well as their software.  Many biologists instead use proprietary
software to prepare their research findings, and many journals officially prefer
Microsoft Word documents as submissions.  The recent discipline-wide adoption of
free software packages such as R and associated document generation systems
\cite{xie2012}, as well as interest in open access publishing, has coincided
with recent rise in use of preprint servers by biologists, supporting Jackson's
claims \cite{xie12}.

>>>>

Submitting to a preprint server is becoming more common in biology, even though
it still involves a minority of papers. The quantitative biology section in
arXiv is experiencing faster growth in submissions than any other field
\cite{cal12}. Also, most scientific journals are preprint-friendly: Nature,
PLOS, BMC, PNAS, Science (mostly) \ref{table:policies}, and all the journals
from Elsevier and Springer. Very recently, the Ecological Society of America and
the Genetics Society of America changed their policies to allow public
preprints. Few scientific publications will not consider a manuscript submitted
to a public preprint server.  Still, a few journals adopt a ``by default''
hostile attitude towards preprints, mostly due to the lack of clear policy of
the publishers, or perhaps because a preprint culture has not developed in
biology and the practice is still considered unusual. As an example,
Wiley-Blackwell, which publishes some leading journals, has no official policy
on the subject \ref{table:policies}.

@@

% Not sure about this paragraph, as much as I love to hate MS Office, I would
% rather keep the thing as a "pro-preprint" paper only and not use it to promote
% LaTeX/R.
% -- PhDP
There are other barriers to adoption of preprint servers by biologists.  The
most notable preprint server, arXiv, is designed to work best with \LaTeX, the
dominant document preparation system among physicists and mathematicians.
Jackson \cite{jackson2002preprints} argues that \LaTeX introduced an
open source mindset to its users, who now freely share their research
findings as well as their software.  Many biologists instead use proprietary
software to prepare their research findings, and many journals officially prefer
Microsoft Word documents as submissions.  The recent discipline-wide adoption of
free software packages such as R and associated document generation systems
\cite{xie2012}, as well as interest in open access publishing, has coincided
with recent rise in use of preprint servers by biologists, supporting Jackson's
claims \cite{xie12}.
